---
title: "m01_v01_churn_predictions"
output: 
    html_document:
        number_sections: true
        #toc: TRUE
        #toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F, fig.width = 13, fig.height = 8)
```

# Company

### The TopBank Company:

A banking services company, with customers mainly in Europe.    
Offers financial products such as bank account, investments and insurance.    
Business Model: banking service through physical and online branches.  
Main product: bank account free of charge, valid for 12 months. After that period, the account must be renewed.  
Bank account revenue per customer:  
  
  - 15% of the client's estimated salary, for clients with estimated income below the average.  
  - 20% of the client's estimated salary, for clients with estimated income above the average.  

### Problem:  

  - The rate of customer cancellation has increased significantly in recent months.  
  - A taxa de cancelamento de clientes aumentou significativamente nos últimos meses.  

### goal:  

  - Reduce customer churn, that is, prevent the customer from canceling the contract and not renewing it for       another 12 months. 
  - Reduzir o churn de clientes, ou seja, evitar que o cliente cancele o contrato e não o renove por mais 12 meses.  

### Deliverable:  

Model performance and results report with the following topics:    
  
  - What is the company's current turnover rate? 
  - Qual é a taxa de rotatividade atual da empresa?
  
  - How does the churn rate vary by month?  
  - Como a taxa de churn varia por mês?  
  
  - How does the model perform to label customers as churns? 
  - Qual é o desempenho do modelo para prever os clientes como churns?  
  
  - - What is the company's revenue, if it prevents customers from entering the contract without canceling it using the model developed?    
  - Qual é a receita da empresa, se ela evita-se que os clientes entrem no cancela-sem o contrato por meio do modelo desenvolvido?  

Action Plan: 

  1. Offer a discount coupon or other financial incentive.  
  2. Which customers could receive the incentive and at what cost, in order to maximize ROI (Return on Investment)?     
  (the sum of the incentives must not exceed $ 10,000.00.)      
  
Plano de Ação : 
  
  1. Oferecer cupom de desconto ou outro incentivo financeiro.  
  2. Quais clientes poderiam receber o incentivo e a que custo, a fim de maximizar o ROI (Return on        Investment)?      
  (a soma dos incentivos não deve exceder $ 10.000,00.)    


# Importing Needed packages

```{r}

library(dplyr)
library(data.table)

```

## Helper Functions

## Reading the data

```{r}

data_raw <- data.table::fread("/home/renato/repos/churn/data/churn.xls")

```

# Descricption of Data

## Rename Columns

```{r}

data_raw <- data_raw %>% janitor::clean_names() 

head(data_raw)

```
## Features Description

**RowNumber -** The number of the row.  
**RowNumber -** O numero de linhas.  

**CustomerID -** Customer's unique identifier.  
**CustomerID -** Identificador único do cliente.  

**Surname -** Customer's surname.  
**Surname -** Sobrenome do Cliente.  

**CreditScore -** Customer's credit score for the consumer market.  
**CreditScore -** Pontuação de crédito do cliente para o mercado consumidor.  

**Geography -** The country where the customer lives.  
**Geography -** O país onde o cliente mora.  

**Gender -** Customer's gender.  
**Gender -** Sexo do Cliente.  

**Age -** Customer's age.  
**Age -** Idade do Cliente.  

**Tenure -** Number of years that the customer was active.  
**Tenure -** O numero de anos que Cliente esteve ativo.  

**Balance -** The amount that the customer has in the bank account.  
**Balance -** Saldo da Conta bancaria.  

**NumOfProducts -** The number of products bought by the customer.  
**NumOfProducts -** O número de produtos comprados pelo cliente.  

**HasCrCard -** Flag that indicates if the customer has a credit card.  
**HasCrCard -** Se o cliente possui cartão de credito.  

**IsActiveMember -** Flag that indicates if the customer has done a bank activity in the last 12 months.  
**IsActiveMember -** Sinalizador que indica se o cliente realizou uma atividade bancária nos últimos 12 meses.  

**EstimateSalary -** Estimate customer's monthly income.  
**EstimateSalary -** Estimativa de renda mensal do Cliente.  

**Exited -** Flag that indicates if the customer is in Churn.  
**Exited -** Indica se cliente cancelou ou nao o contrato.  


## Data Dimensions

```{r}

print(paste("Number of Rows: " ,nrow(data_raw)))
print(paste("Number of Cols: " ,ncol(data_raw)))

```

## Data Types

```{r}

glimpse(data_raw)

```

## Checking NA

```{r}

colSums(is.na(data_raw))

```

- The data set has no missing values.  
- O conjunto de dados não possui valores ausentes.  

## Change Types

```{r}

data_raw <- data_raw %>% 
  dplyr::mutate_if(is.character, as.factor)

```

## Descriptive Statistics

```{r}

# selecting only numeric features
num_attributes <- data_raw %>% 
  purrr::keep(is.numeric)

# selecting only categorical features
cat_attributes <- data_raw %>% 
  purrr::keep(is.factor)

```

### Numeric Attributes

```{r}

# Central Tendency  - mean , median
num_mean <- as.data.frame( t(lapply(num_attributes, mean)))

num_median <- as.data.frame( t(lapply(num_attributes, median)))

# dispersion - std, min, max, range, skew, kurtosis
num_std <- as.data.frame( t(lapply(num_attributes, sd)))

num_min <- as.data.frame( t(lapply(num_attributes, min)))

num_max <- as.data.frame( t(lapply(num_attributes, max)))

num_skew <- as.data.frame( t(lapply(num_attributes, e1071::skewness)))

num_kurt <- as.data.frame( t(lapply(num_attributes, e1071::kurtosis)))

table_desc <- t(bind_rows(num_min,num_max,num_mean,num_median,num_std,num_skew,num_kurt))

table_desc<- as.data.frame(table_desc)

names(table_desc) <- c("min","max","mean","median","std","skew", "kurtosis")

knitr::kable(table_desc, digits = 4) %>% 
  kableExtra::kable_styling(full_width = F, bootstrap_options = c("striped", "hover", "condesend", "responsive"),html_font = "Cambria")

```

```{r, fig.height= 10}

num_attributes %>%
  purrr::keep(is.numeric) %>% 
  tidyr::gather() %>% 
  ggplot2::ggplot(ggplot2::aes(value)) +
    ggplot2::facet_wrap(~ key, scales = "free") +
    ggplot2::geom_histogram(col= "black", fill="steelblue", bins = 25)+
    ggplot2::scale_y_continuous(labels = function(x) format(x, scientific = FALSE))+
  ggplot2::labs(title = "Distribution of numerical variables")+
  ggplot2::scale_x_continuous(labels = function(x) format(x, scientific = FALSE))+
  ggplot2::theme(plot.title = ggplot2::element_text(hjust = 0.5, size = 18))

```
  


```{r}

data_raw %>%
  mutate(exited = as.factor(exited)) %>% 
  dplyr::count(exited) %>% 
  mutate(prop = round(n/sum(n)*100,2)) %>% 
  ggplot2::ggplot(ggplot2::aes(x = exited, y = prop, color = exited)) +
  ggsci::scale_color_jco() +
  ggplot2::geom_segment(ggplot2::aes(xend = exited, yend = 0), show.legend = F) +
  ggplot2::geom_point(ggplot2::aes(size = prop), show.legend = F) +
  ggplot2::geom_label(ggplot2::aes(label = paste0(prop,"% / n = ",n)),
             fill = "white", 
             hjust = "inward",
             show.legend = F) +
  ggplot2::labs(y = "%",
       x = "Exited") +
  ggplot2::coord_flip() +
  ggplot2::theme_minimal()

```

The company's current cancellation rate is 20%.  
A taxa de cancelamento atual da empresa é de 20%.  


### Categorical Attributes

```{r}

apply(cat_attributes, 2, function(x) length(unique(x)))

```

```{r}

ggpubr::ggarrange(
data_raw %>%
  dplyr::count(gender) %>% 
  mutate(prop = round(n/sum(n)*100,2)) %>% 
  ggplot2::ggplot(ggplot2::aes(x = gender, y = prop, color = gender)) +
  ggsci::scale_color_jco() +
  ggplot2::geom_segment(ggplot2::aes(xend = gender, yend = 0), show.legend = F) +
  ggplot2::geom_point(ggplot2::aes(size = prop), show.legend = F) +
  ggplot2::geom_label(ggplot2::aes(label = paste0(prop,"% / n = ",n)),
             fill = "white", 
             hjust = "inward",
             show.legend = F) +
  ggplot2::labs(y = "%",
       x = "Gender", title = "Gender") +
  ggplot2::coord_flip() +
  ggplot2::theme_minimal()+
  ggplot2::theme(plot.title = ggplot2::element_text(hjust = 0.5, size = 18)),

data_raw %>%
  dplyr::count(geography) %>% 
  mutate(prop = round(n/sum(n)*100,2)) %>% 
  ggplot2::ggplot(ggplot2::aes(x = geography, y = prop, color = geography)) +
  ggsci::scale_color_jco() +
  ggplot2::geom_segment(ggplot2::aes(xend = geography, yend = 0), show.legend = F) +
  ggplot2::geom_point(ggplot2::aes(size = prop), show.legend = F) +
  ggplot2::geom_label(ggplot2::aes(label = paste0(prop,"% / n = ",n)),
             fill = "white", 
             hjust = "inward",
             show.legend = F) +
  ggplot2::labs(y = "%",
       x = "Geography", title = "Geography") +
  ggplot2::coord_flip() +
  ggplot2::theme_minimal()+
  ggplot2::theme(plot.title = ggplot2::element_text(hjust = 0.5, size = 18)),

ncol = 2)

```

Conclusions:  
Conclusões:  

- The average age of customers is 38 years.  
- A média de idade dos clientes esta em 38 anos.  

- Balance has an almost normal distribution, with most customers between 100,000.00 and 130,000.00 euros in their bank account.  
- Balance possui uma distribuição quase normal, tendo na sua maioria clientes entre 100,000.00  a 130,000.00 euros em sua conta bancaria.  

- Most customers have an average score of 400 to 700.  
- A maioria dos clientes possui um score médio de 400 a 700.  

- Customers have an average of 100,000.00 euros per month.  
- Os clientes possuem em média 100,000.00 euros por mês.  

- Most customers have a credit card.  
- A maioria do clientes possui cartão de credito.  

- Customers who have and have not done banking in the past 12 months are almost balanced.  
- Os clientes que realizaram e nao realizaram um operação bancaria no ultimos 12 meses é quase equilibrado.  

- Most customers own at least one bank product.  
- A maioria dos clientes possuem pelo menos um produto do banco.  

- The average number of customers who are active in the bank is 5 years.  
- A média de clientes que estão ativos no banco é de 5 anos.

- Only 20% is healthy churn.  
- Apenas 20% é são churn.  

- 55% of customers are male.  
- 55% dos clientes são masculino.  

- 50% of customers are from France.  
- 50% dos clientes são da frança.  

```{r}

num_attributes %>% 
  filter(balance == 0) %>% 
  count()
  

```



